---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Case Study: How Does a Bike-Share Navigate Speedy Success
\
\


# Introduction

### 1. Scenario
```{r - Tableau presentation1, echo=FALSE, fig.align='center', fig.show='hold', out.width="100%"}
knitr::include_graphics(
  c("~/R/Cyclistic/images/02.png")
)
```
\
\

### 2. Stakeholders

- Director of marketing, Lily Moreno, my manager: She is responsible for marketing campaigns to promote the bike-share program. \
- Marketing analytics team: A team of data analysts, responsible for collecting, analyzing, and reporting data for the market strategy. I am a junior data analyst of this team.\
- Executive team: Will decide whether to approve the recommended marketing program.\
\
\

### 3. Data Analysis Process

To answer the business question, we will follow the data analysis process: **ask, prepare, process, analyze, share**, and **act.**\
\

#### 3.1 Ask

Three questions will guide the marketing program:\

1. How do annual members and casual riders use Cyclistic bikes differently?\
2. Why would casual riders buy Cyclistic annual memberships?\
3. How can Cyclistic use digital media to influence casual riders to become members?\

The marketing director has assigned me to answer the first question?\
\
**`How do annual members and casual riders use Cyclistic bikes differently?`**\
\

#### 3.2 Prepare

For the data analysis we will use trip data from previous 12 months from this [source](https://divvy-tripdata.s3.amazonaws.com/index.html). (Note: The data has been made available by Motivate Inernational Inc. under this [license](https://ride.divvybikes.com/data-license-agreement).)\
\

#### 3.3 Process

**a) Choose the right tools:**\

The use of spreadsheets can be ruled out due to the large data size, leaving us with SQL or R as best choices.

Approach 1: \
- *Excel* (Pivot and Graphs): For an ad-hoc analysis with a single sample file.\
- *SQL Server* Management Studio: For data import and cleaning\
- *Tableau desktop*: for data exploration, visualization, and reporting\

Approach 2:\
- *R: RStudio/ RMarkdown*: for a more in-depth analysis and for reporting\
- *GitHub*: reporting and code sharing\
\

**b) Data integrity**\

Data integrity is to ensure that the data sample is sufficient, has no bias, represents the population, is accurate and trustworthy.\
In this project the data sample represents the total population of one year, is therefore not biased, and the data are from first party.\
\

 **c) Cleaning the data**\
 
The provided raw data are already in relative clean condition. Nevertheless, the following cleaning steps are still required:\

- removing missing data\
- removing irrelevant data \
- handling datetime\
- removing duplicates\
- computing new variables for further analysis\
- removing outliers\
\

#### 3.4 Analyze

In the analysis process the cleaned dataset is aggregated in various ways to answer questions about trends per year, week, day, bike type and location. The results are summarized in visualization.\
\

#### 3.5 Share

All findings are visualized and reported in a Rmarkdown document and shared on GitHub. The Tableau story line and interactive dashboards are available on Tableau Public Server [Link](https://public.tableau.com/app/profile/hartmut.schaefer/viz/Bike-ShareCaseStudy-GoogleDataAnalyticsCapStoneProject/Cyclistics-StoryPoint)\
\

#### 3.6 Act

Recommendations are stated in each report.\
\
