---
title: "Project in SQL and Tableau"
output: 
  html_document:
     code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


# Data Cleaning in SQL
\

In this RMarkdown file the SQL code for data cleaning is documented. The SQL tool used for this case study is Microsoft SQL Server Management Studio v.18.12.1.\
\
This RMarkdown file is setup to connect to my localhost SQL Server. All SQL code chunks will be then executable.\
\
Note: An issue left for further investigation: 
The SQL tables used here cannot be fetched into an R dataframe due to mixed variable lengths of some columns. According to [StackOverflow](https://stackoverflow.com/questions/45001152/r-dbi-odbc-error-nanodbc-nanodbc-cpp3110-07009-microsoftodbc-driver-13-fo) the variables in the SQL table must be sorted from short to long before fetching.  For our purpose of documentation fetching is not required. However, if an SQL table has to be further processed in R this issue has to be addressed.
\

### Prerequisites
\

#### Install and load packages:

Install packages "ODBC" and "DBI" for database communication:
```{r load packages, message=FALSE, warning=FALSE}
library(odbc)
library(DBI)
library(tidyverse)
```


#### Connect to SQL Server:
```{r set up connection}
con <- dbConnect(odbc(), Driver = "SQL Server",
                 server = "localhost\\SQLEXPRESS",
                 Database = "Cyclistic_2022",
                 Trusted_Connection = "True")
```
\



## 1. Importing and merging raw data

### 1.1 Importing data

The following datasets were used:[Link for dataset download:](https://divvy-tripdata.s3.amazonaws.com/index.html) \

Data set: '202201-divvy-tripdata.csv' (data for one month Jan. 2022)\
Data set: '202202-divvy-tripdata.csv' (data for one month Feb. 2022)\
Data set: '202203-divvy-tripdata.csv' (data for one month Mar. 2022)\
Data set: '202204-divvy-tripdata.csv' (data for one month Apr. 2022)\
Data set: '202205-divvy-tripdata.csv' (data for one month May. 2022)\
Data set: '202206-divvy-tripdata.csv' (data for one month Jun. 2022)\
Data set: '202207-divvy-tripdata.csv' (data for one month Jul. 2022)\
Data set: '202208-divvy-tripdata.csv' (data for one month Aug. 2022)\
Data set: '202209-divvy-tripdata.csv' (data for one month Sep. 2022)\
Data set: '202210-divvy-tripdata.csv' (data for one month Oct. 2022)\
Data set: '202211-divvy-tripdata.csv' (data for one month Nov. 2022)\
Data set: '202212-divvy-tripdata.csv' (data for one month Dec. 2022)\
\

All raw data files are imported to SQL Server; First, before the data import the PC/Mac System Clock must be changed to "Central Standard Time (Chicago)". In this way the datetime variables in the datasets are assigned to the local time zone. Second, during the import the datetime format must be changed to "datetimeoffset(7)". In this way the local daylight saving time offset is registered correctly.\
\



**Analysis: Customer behavior differences by bike type**
\

```{r - Tableau presentation5, echo=FALSE, fig.align='center', fig.show='hold', out.width="100%"}
knitr::include_graphics(
  c("~/R/Cyclistic/images/05.png")
)
```
\


