---
title: "References"
output: 
  html_document:
    toc: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



The Google Data Analytics course provided the basic and broad foundation to the field of data analysis.  To challenge the case study project, however, I needed some more training on the tools. Below are some recommendations for further tutoring and reading. I also used tons of other sources on the open net but mainly `StackOverflow`.\
\
\



1. [Google Data Analytics Professional Certificate, Coursera](https://www.coursera.org/professional-certificates/google-data-analytics?utm_medium=sem&utm_source=bg&utm_campaign=B2C_NAMER_google-data-analytics_google_FTCOF_professional-certificates_country-CA&campaignid=662140813&adgroupid=1255642923849196&device=c&keyword=google%20data%20analytics%20certification%20coursera&matchtype=e&network=o&devicemodel=&adposition=&creativeid=&hide_mobile_promo&msclkid=761b632d0cbd1554c7b5070b07180e53&utm_term=google%20data%20analytics%20certification%20coursera&utm_content=Certificate%20Related)\
\
2. [SQL Server Tutorial, Youtube - PRAGIM Technologies, India]\
\
3. [Tableau Tutorial (1-87), Youtube - tutorialspoint, India]\
\
4. [R for Data Science, O'Reilly, Hadley Wickham](https://r4ds.had.co.nz/index.html)\
\
5. [Let's Git started, Jenny Bryan, Jim Hester, UBC](https://happygitwithr.com/index.html)\
\
6. [Toronto bike-Share Data Analysis, Shirly Zhang, Medium](https://shirleyzhang2.medium.com/toronto-bike-share-data-analysis-943d5810c717)\
\






```{r load packages, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(here)
```
\

```{r eval=FALSE, include=FALSE}
reference_link <- here("data", "References.csv")
references_list <- read_csv(reference_link)
```

```{r eval=FALSE, include=FALSE}
# --------------------------------------------------------------- print as table
knitr::kable(references_list[1:6 , ])
```

